\Title

### 2021-03-16 latex/paper/main.tex
First draft of the problem formulation section

%---------- Theorems ----------%
\newtheorem{thm}{Theorem}% Uncomment: reset numbering at each chapter
%\newtheorem{prop}[thm]{Proposition} % Propositions depend on Theorem number
\newtheorem{lemma}[thm]{Lemma} % Same with lemmas
\newtheorem{defn}[thm]{Definition} % Definitions depend on theorem number
\newtheorem{assm}{Assumption} % Assumptions do not reset

@@

%/========== Problem Formulation ==========/%
\section{Problem Formulation}\label{sec:problem-formulation}
TODO: motivate injection/dissipation for energy regulation in an acrobot. Use
this to lead into VNHCs. e.g. represent torso as chain of a swing, pivot as
seat, and legs as human legs. Replicate fig3.3 from thesis but for acrobot: take
(qu,pu)-plane and show how a person moves their legs wrt theta, qu, or pu. Leave
the gymnast model until the end, or ignore it entirely.

Alternatively, ff human movement journal says legs move wrt qu, we can use that
as a foundation.

@@

This system has inertia matrix 
\(D\), potential function \(P\) (with respect to
the horizontal bar), and input matrix \(B\) given as follows:
\begin{align}\label{eqn:acrobot-inertia}
    D_s(q) &= \begin{bmatrix}
        ml^2\left(3+2\cos(q_a)\right) & 
        ml^2\left(1+\cos(q_a)\right) \\
        ml^2\left(1+\cos(q_a)\right) &
        ml^2
    \end{bmatrix} 
    , \\
    \label{eqn:acrobot-potential}
    P_s(q) &= -mgl\left(2\cos(q_u)+\cos(q_u+q_a)\right)
    , \\
    \label{eqn:acrobot-B}
    B &= [0;1]
    .
\end{align}
%/========== Simulation ==========/%
\section{Simulation Results}\label{sec:simulations}

>>>>

%---------- Theorems ----------%
\newtheorem{thm}{Theorem}% Uncomment: reset numbering at each chapter
\newtheorem{prop}[thm]{Proposition} % Propositions depend on Theorem number
\newtheorem{lemma}[thm]{Lemma} % Same with lemmas
\newtheorem{defn}[thm]{Definition} % Definitions depend on theorem number
\newtheorem{assm}{Assumption} % Assumptions do not reset

@@

%/========== Problem Formulation ==========/%
\section{Problem Formulation}\label{sec:problem-formulation}
%TODO: motivate injection/dissipation for energy regulation in an acrobot. Use
%this to lead into VNHCs. e.g. represent torso as chain of a swing, pivot as
%seat, and legs as human legs. Replicate fig3.3 from thesis but for acrobot: take
%(qu,pu)-plane and show how a person moves their legs wrt theta, qu, or pu. Leave
%the gymnast model until the end, or ignore it entirely.
%
%Alternatively, ff human movement journal says legs move wrt qu, we can use that
%as a foundation.
%Ideas: Human movement says appropriate regulation of pendulum allows energy to
%be pumped into system. Pendulum length rapidly increased on downward swing, then
%rapidly decreased on upward swing (I think these mean backward and forward).
%They plot pendulum length rdot/r vs "body angle" theta to determine energy
%injection, which hints at some relationship between (r,rdot) and (theta,thetad).
%Indeed, the peak value of rdot/r did not occur at a fixed value of theta, nor
%once anyone reached a particular speed; it occured at the same thetadot/theta for all
%participants of the study. This perfectly motivates VNHCs for robots, because we
%can shape (r,rtdot) as functions of (theta,thetad).
%Start motivation from energy injection - how do humans inject energy? How does
%acrobot model gymnast? use human movement to motivate tracking functions of both
%position and velocity rather than time.

In gymnastics terminology, a ``giant" is the motion a gymnast performs to
achieve full rotations around a horizontal bar \cite{usagym_giant}. 
A gymnast will begin (at rest) by hanging on a bar, then swing their legs
appropriately to gain energy over time.
The authors of \cite{pendulum_length_giant_gymnastics} modelled the gymnast as a
variable length pendulum, and studied how the pendulum's length changes as a
function of the orientation of a gymnast's joints.
Labeling the pendulum length by \(r\) and the gymnast's body orientation
by \(\theta\), they discovered experimentally that the value \(\dot{r}/r\) has
the highest impact on the magintude of energy injection; they also discovered
that the peak value of \(\dot{r}/r\) occured at the same value of
\(\dot{\theta}/\theta\) for all gymnasts.
In other words, gymnasts were moving their legs as a function of their body's
angle and their angular velocity around the bar, not as functions of time.

Typically as control theorists we like our system's actuatiors to track a fixed
trajectory over time; but if we were attempting design a gymnastics robot that
could replicate human behaviour, the results of
\cite{pendulum_length_giant_gymnastics} indicate that we should not use this
standard approach.
Instead, we should have our robot's legs follow a function of its
body angle and angular velocity.
That is exactly what we set out to do in this article.

The simplest way to model the gymnast is as a variable-length pendulum, but a
more realistic model is that of a two-link acrobot (Figure \ref{fig:acrobot}).
The acrobot's first link represents the gymnast's torso, the bottom link represents
their legs, and it is actuated at the center joint (the hips).

\begin{figure}
    \centering
    %\includegraphics[width=\textwidth]{images/acrobot.png}
    \caption{The two-link acrobot.}
    \label{fig:acrobot}
\end{figure}

In this article, our goal is to design a leg controller for the acrobot so that
the resulting motion mimics that of a human performing a giant.
Previous approaches to solve this problem used
trajectory tracking, set-point tracking, or energy
pumping regulation (see
\cite{energy_pumping_robotic_swinging,swingup_giant_acrobot,dynamical_servo_acrobot_vc,control_giant_two_link_gymnastic_robot}
).
While they all result in the acrobot performing rotations on the bar, none of
the are particularly human-like.
The most realistic controller to date was designed by Wang \cite{xingbo_thesis}; it
forces the hips to track a function of the torso angle.
However, a truly human-like controller would force
the hips to track a function involving the torso velocity.
These kinds of controllers are known as ``virtual nonholonomic constraints", and
they have been used for human-robot interaction
\cite{vnhc_human_robot_cooperation,psd_based_vnhc_redundant_manipulator,haptic_vnhc},
error-reduction on time-delayed systems \cite{vnhc_time_delay_teleop},
and they have shown marked improvements to the field of bipedal locomotion 
\cite{nhvc_dynamic_walking,
hybrid_zero_dynamics_bipedal_nhvcs,output_nhvc_bipedal_control}.
Indeed, virtual nonholonomic constraints produce more robust bipedal motion than
controllers which do not depend on velocity \cite{nhvc_incline_walking}.
In particular, virtual nonholonomic constraints may be capable of injecting and
dissipating energy from a system in a robust manner, all while producing
realistic biological motion. 
Hence, we will design one of these to achieve stable giant motion on the
acrobot.

Here then is the problem we will solve: design a virtual nonholonomic constraint
for the acrobot, so that the acrobot provably gains energy over time
(as per Definition \ref{defn:energy-gain}).

\begin{defn}\label{defn:energy-gain}
    Let \(\mathcal{Q}\) be an
    \(n\)-dimensional generalized cylinder.
    Let \(f : \mathcal{Q} \rightarrow \mathcal{Q}\times\R^n\) be a smoth vector
    field and let \(D \subset M\) be open.
    The system described by \(\dot{x} = f(x)\) 
    \textit{gains energy on \(D\)} if, 
    for all compact sets \(K \subset D\) and for almost every initial
    condition \(x(0) \in K\), there exists \(T > 0\) such
    that \(x(t) \notin K \, (\forall t > T)\).
    The system \textit{loses energy on \(D\)} if it gains energy in
    negative-time.
\end{defn}

@@

This system has inertia matrix 
\(M\), potential function \(V\) (with respect to
the horizontal bar), and input matrix \(B\) given as follows:
\begin{align}\label{eqn:acrobot-inertia}
    M(q) &= \begin{bmatrix}
        ml^2\left(3+2\cos(q_a)\right) & 
        ml^2\left(1+\cos(q_a)\right) \\
        ml^2\left(1+\cos(q_a)\right) &
        ml^2
    \end{bmatrix} 
    , \\
    \label{eqn:acrobot-potential}
    V(q) &= -mgl\left(2\cos(q_u)+\cos(q_u+q_a)\right)
    , \\
    \label{eqn:acrobot-B}
    B &= [0;1]
    .
\end{align}
The conjugate of momenta is \(p = (p_u,p_a) = M(q)\dot{q}\).
The dynamics of the acrobot in \((q,p)\) coordinates are given in
~\eqref{eqn:acrobot-hamiltonian}, where
for shorthand, we write \(c_u := \cos(q_u)\), \(c_a := \cos(q_a)\), and 
\(c_{ua} := \cos(q_u + q_a)\); likewise, \(s_u := \sin(q_u)\), 
\(s_a := \sin(q_a)\), and \(s_{ua} := \sin(q_u + q_a)\).
\begin{align}\label{eqn:acrobot-hamiltonian}
    \mathcal{H}(q,p) &= \frac{1}{2}p\tpose \Minv(q) p -
    mgl\left(2 c_u + c_{ua}\right)
    , \\
     &\begin{cases}
        \dot{q} = \Minv(q) p 
        ,\\
        \dot{p}_u = -mgl\left(2s_u + s_{ua}\right) 
        ,\\
        \dot{p}_a =-\frac{1}{2}p\tpose \nabla_{q_a}\Minv(q) p
        - mgl s_{ua} + \tau,
    \end{cases} \nonumber
\end{align}
The control input is a force \(\tau \in \R\) affecting only the dynamics of
\(p_a\), representing a torque acting on the hip joint.
This means \((q,p)\) are simply actuated coordinates inside the phase space
\(\mathcal{Q} \times \mathcal{P}\) where
\(\mathcal{Q} = \mathcal{Q}_u \times \mathcal{Q}_a 
:= \Sone \times \Sone\), and
\(\mathcal{P} = \mathcal{P}_u \times \mathcal{P}_a
:= \R \times \R\).
Hence, we can apply the theory of VNHCs from Section \ref{sec:vnhc} to design an
energy-injecting controller.
Since we need our VNHC to be regular, the following proposition will prove
useful.
\begin{prop}\label{prop:acrobot-fpu-regular}
    A relation \(h(q,p) = q_a - f(p_u) = 0\) 
    with \(f \in C^2\left(\mathcal{P}_u; \mathcal{Q}_a\right)\) is a regular
    VNHC of order 1 for the simple acrobot.
\end{prop}
\begin{proof}
    The regularity matrix for this VNHC evaluates to
    \(\frac{(1+c_a)\partial_{q_u}f(q_u,p_u) + (3+2c_a)}{ml^2(2-c_a^2)}\).
    Since \(\partial_{q_u} f = 0\), this matrix is strictly positive for all values
    of \(q_a\), and hence is full rank everywhere on the constraint manifold.
\end{proof}

%/========== Simulation ==========/%
\section{Simulation Results}\label{sec:simulations}
